---
title: "Isotope Samples"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Set up

Go through this document one chunk at a time to insert new isotope data into the database. First, connect to the database, load libraries, and load all of the functions in [rodm2](https://github.com/khondula/rodm2).

```{r}
library(RPostgreSQL)
library(readr)
library(dplyr)
library(tidyr)
library(uuid)
# source("load_rodm2_functions.R")
```

# Upload and prepare data

Upload the spreadsheet with your new sample data. 

```{r}
new_samples <- read_csv("data/Copy of Isotopes_4Kelly.csv")
```

Make a column `sample_code` that will be unique sample IDs. 

```{r}
new_samples$sample_code <- new_samples$sample_name
```

Then make sure that the sample codes don't already exist in the database by loading the samplingfeatures table and comparing the samplingfeaturecode column to the new_samples sample code column. (If they do exist and new results are being added, skip ahead to **Insert sample results data**.)

```{r}
samplingfeatures <- dbReadTable(db, c("odm2", "samplingfeatures"))

data.frame(sample_code = new_samples$sample_code,
           in_db = new_samples$sample_code %in% samplingfeatures$samplingfeaturecode) %>%
  dplyr::filter(in_db) %>% arrange(sample_code)
```

Then also make sure that parent sampling features exist (i.e. the sites that the samples were collected from). 

Sites not in database: 

```{r}
data.frame(sample_site = new_samples$origin_site,
           in_db = new_samples$origin_site %in% samplingfeatures$samplingfeaturecode) %>%
  dplyr::filter(!in_db) %>% arrange(sample_site)
```

DO, DW, JA or JB, JB or JA, ML, P1, TM

There needs to be a defined method in the methods table about how samples were collected. Make sure that sample collection method exists and note the **methodcode** or make a new one with the `db_add_method()` function. 

```{r}
dbReadTable(db, c("odm2", "methods"))
methodcode = "waterisotopefield"
```


```{r, eval=FALSE}
db_add_method(methodname = "water isotope sample collection",
              methodcode = "waterisotopesample",
              methodtypecv = "Specimen collection",
              methoddescription = "Collect water sample without headspace for water isotopes")

```

There needs to be a correctly formatted date and time of the sample collection action. Format the date and time. If no time is given, assume midnight (to know that it isn't a specific time). Also define the UTC offset variable. 

```{r}
utcoffset <- format(Sys.time(), "%z")
utcoffset <- as.integer(substr(utcoffset, 1,3))
new_samples$datetime <- strptime(paste(new_samples$origin_sample_date, "12:00:00"),
         format = "%Y-%m-%d %H:%M:%S", tz = "")
```

Make any new annotations to use based on notes about samples using `db_add_anotation()`

```{r, eval=FALSE}
dbReadTable(db, c("odm2", "cv_annotationtype"))
db_add_annotation(annotationtypecv = "Specimen annotation",
                  annotationtext = "collected less than 24 hours after rain")
db_add_annotation(annotationtypecv = "Specimen annotation",
                  annotationtext = "collected before forecast rain")
```

Add a column of unique UUIDs.

```{r}
new_samples$new_samples_uuid <- sapply(1:nrow(new_samples), UUIDgenerate)
```

Make a descriptive name for each sample (not required)

```{r}
new_samples$sample_description <- paste("Water isotope sample",
                                        new_samples$sample_name,
                                        "collected from",
                                        new_samples$sample_category,
                                        "at site",
                                        new_samples$origin_site,
                                        "on",
                                        new_samples$origin_sample_date
                                        )
```

Define some necessary parameters for the specimens table.

```{r}
specimentypecv = "Grab"
specimenmediumcv = "Liquid aqueous"
isfieldspecimen = TRUE
```

# Insert sample collection actions.

Load function `db_insert_samples()` if not already loaded, and run it over each line in the new samples data frame. This creates a new specimen in the sampling feature table for each sample, relates them each to the site they were collected at (`origin_site`), and makes a feature action of when the sample was collected. 

```{r, eval=FALSE}
source("db_insert_samples.R")
sapply(2:nrow(new_samples), function(x) db_insert_samples(x))
```

Then add any annotations about samples or sample collection. 

```{r}
new_annotations <- new_samples %>% 
  dplyr::select(sample_code, notes) %>%
  dplyr::filter(!is.na(notes)) %>%
  rename("samplingfeaturecode" = sample_code,
         "annotation" = notes)

new_annotations$annotation <- gsub(" <", "less than", new_annotations$annotation)

# make sure the annotations exist in the database
new_annotations$annotation %in% 
  dbGetQuery(db, "SELECT annotationtext from odm2.annotations")$annotationtext

```

insert annotations

```{r, eval=FALSE}
sapply(1:nrow(new_annotations), function(x) insert_samplingfeature_annotations(x))
```


# Insert sample results data

After the samples exist in the database and have sample collection feature actions, results are associated with new sample analysis feature actions. 

Check that the necssary methods and variables exist in the database and make a note of the code names. 

```{r, eval=FALSE}
db_add_method(methodtypecv = "Specimen analysis",
  methodcode = "waterisotopelab",
  methodname = "Picarro L1102-i",
  methoddescription = "D-18O and  D-D ratios measured using a cavity ringdown spectroscopic liquid water laser isotope analyzer (Picarro L1102-i) in the Virginia Tech Water Quality Lab housed within the Biological Systems Engineering Department.")
```

add new variables for d180 and dD

```{r, eval=FALSE}
# dbReadTable(db, c("odm2", "cv_variabletype"))$name
db_add_variable(variabletypecv = "Stable isotopes",
                variablecode = "delta_18OOfH2O",
                variabledefinition = "Isotope 18O of water",
                variablenamecv = "delta-18O of H2O")
db_add_variable(variabletypecv = "Stable isotopes",
                variablecode = "delta_DOfH2O",
                variabledefinition = "hydrogen isotopes of water",
                variablenamecv = "delta-D of H2O")
```

Prepare a data frame of new sample results called `new_sample_results` with columns sample_code, `d 18O`, `d D`, and `origin_sample_date. Only have rows with data to upload. 

```{r}
new_sample_results <- new_samples %>%
  dplyr::select(sample_code, `d 18O`, `d D`, origin_sample_date)
new_sample_results <- new_sample_results %>% dplyr::filter(!is.na(`d 18O`))
```

Format time correctly. Need a time of the sample analysis action. The data value can be associated with the sample collection datetime so keep that in there too. 

```{r}
new_sample_results$collection_datetime <- 
  strptime(paste(new_sample_results$origin_sample_date, "12:00:00"),
         format = "%Y-%m-%d %H:%M:%S", tz = "")

new_sample_results$analysis_datetime <- "2018-04-23 12:00:00"
```

Specify the units.

```{r}
unitsabbreviation = "per mil VSMOW"
```

Specify the variable codes. 

```{r}
variablecode1 = "delta_18OOfH2O"
variablecode2 = "delta_DOfH2O"
```

Specify other necessary metadata parameters. 

```{r}
utcoffset <- format(Sys.time(), "%z")
utcoffset <- as.integer(substr(utcoffset, 1,3))
actiontypecv = "Specimen analysis"
resulttypecv = "Measurement"
censorcodecv = "Not censored"
qualitycodecv = "Unknown"
aggregationstatisticcv = "Sporadic"
timeaggregationinterval = 1 # assume 1 minute
timeaggregationintervalunitsid = dbGetQuery(db, "SELECT unitsid FROM odm2.units WHERE unitsname = 'Minute'") # minute
processlinglevelid = 2 # definition = "Quality controlled data" # processinglevelcode = 1
valuecount = 1
sampledmediumcv = "Liquid aqueous"
```

Load function to insert sample results and run function over all rows in dataframe. This is where the magic happens!

```{r, eval=FALSE}
sapply(1:nrow(new_sample_results),
       function(x) db_insert_samle_results(x))
```

