---
title: "Connect to choptank-db"
output:
  html_notebook:
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(magrittr)
```

Make sure you have

* installed the `RPostgreSQL` library with `install.packages("RPostgreSQL")`.
* saved the database password as `.pgpass` in your working directory.
* connected to VPN or are using rstudio.sesync.org

```{r, message=FALSE}
library(RPostgreSQL)
```

Read in password if you have it saved in your working directory:
```{r pw, message=FALSE, results='hide'}
password <- scan(".pgpass", what="")
```
Or create a prompt to enter it in the Console window:
```{r}
password <- readline(prompt="Enter password: ")
```


Create connection object called `db`. These guides will always use the name `db` for the connection object. All of the functions that send and receive information from the database will have this as the first argument. 

```{r}
db <- dbConnect(PostgreSQL(), 
                host = "sesync-postgis01.research.sesync.org",
                dbname = "choptank", 
                user = "palmergroup",
                password = password)
```

If you want to read a whole table from the database into your R environment, you can use `dbReadTable()`. 
```{r}
samplingfeatures <- dbReadTable(db, c("odm2", "samplingfeatures"))
```

Disconnect at the end of your session

```{r}
dbDisconnect(db)
```
